---
title: "simualtions"
output:
  workflowr::wflow_html:
    code_folding: show
    toc: no
  html_notebook: default
  html_document:
    df_print: paged
    toc: no
editor_options:
  chunk_output_type: inline
---

# Introduction

## Introduction

We seek to quantify the enrichment of annotation parameters using multivariate or univariate input. To run the software [TORUS](https://github.com/xqwen/torus) we use a [pipeline](https://github.com/cumc/bioworkflows/blob/master/fine-mapping/gwas_enrichment.ipynb) available using the workflow manager SOS and run the following two lines of code. 


```{r,eval=F,echo=T}
sos run fine-mapping/gwas_enrichment.ipynb range2var_annotation --cwd $work_dir --annotation_dir $anno_dir --z-score $z_file --single-annot $single
sos run fine-mapping/gwas_enrichment.ipynb enrichment --cwd $work_dir --annotation_dir $anno_dir --z-score $z_file --single-annot $single --blocks $blk --snps $snps
```

We place the results of the Torus pipeline in `mvp_complete_torus` where `mash` and `original` refer to the results using univariate and multivariate summary stats respectively.

Here, we plot the meta analysis results:


###HDL

```{r}
library('ggplot2') 
library("tidyverse")


mash.hdl=read.csv(("~/Downloads/mvp_complete_torus//mash/hdl.zscore.torus.merged.csv"))
mash.hdl=mash.hdl[-which(mash.hdl$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]
###Orignal
raw.hdl=read.csv(("~/Downloads/mvp_complete_torus/original/original_z_mvp.hdl.zscore.torus.merged.csv"))
raw.hdl=raw.hdl[-which(raw.hdl$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]


df1=mash.hdl[,c(1:4)]
colnames(df1)=c("Outcome","log2OR","Lower","Upper")
df1$se=(df1$log2OR-df1$Lower)/1.96

df2=raw.hdl[,c(1:4)]

colnames(df2)=c("Outcome","log2OR","Lower","Upper")
df2$se=(df2$log2OR-df2$Lower)/1.96

# add a group column
df1$group <- "mash"
# create a second dataset, similar format to first

# different group
df2$group <- "univariate"
# and we adjust the values a bit, so it will look different in the plot
df2[,c("log2OR","Lower","Upper")] 

# combine the two datasets                      
df = rbind(df1,df2)


z=(df1$log2OR-df2$log2OR)/sqrt(df1$se^2+df2$se^2)
p=2*pnorm(-1*abs(z))




dotCOLS = c("#a6d8f0","#f9b282")
barCOLS = c("#008fd5","#de6b35")
dotCOLS = c("#a6d8f0","gray80")
barCOLS = c("#008fd5","gray80")

df$p=c(p,p)
df$fp = format.pval(df$p,digits=1)

p = df %>% ggplot(aes(y=reorder(Outcome, desc(p)), x=log2OR, xmin=Lower, xmax=Upper, col=group, fill=group)) +
    geom_point(size=3, position=position_dodge(width = 0.5)) +
    geom_errorbarh(height=0, size=1, position=position_dodge(width = 0.5)) +
    geom_vline(xintercept=0, lty=2) +
    # geom_point(size=3, shape=21, colour="white", stroke = 0.5, position=position_dodge(width = 0.5)) +
    scale_fill_manual(values=barCOLS)+
    scale_color_manual(values=dotCOLS)+
    scale_y_discrete(name="Annotation") +
    scale_x_continuous(name="Log2 Odds ratio in HDL", limits = c(-5, 12), breaks=c(-5,0,5,10), labels=c(-5,0,5,"P-value"), position='top') +
    theme_minimal() +
    theme(legend.position=c(0.05,0.05), legend.justification=c(0,0), legend.title=element_blank()) +
    geom_text(aes(x=10, y=Outcome, label=fp), hjust=0.5, color='black')

#pdf("~/Dropbox/hdl_torus.pdf")
p
#dev.off()

```

###LDL
```{r}

rm(list=ls())




library('ggplot2') 

mash.ldl=read.csv(("~/Downloads/mvp_complete_torus//mash/ldl.zscore.torus.merged.csv"))
mash.ldl=mash.ldl[-which(mash.ldl$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]
###Orignal
raw.ldl=read.csv(("~/Downloads/mvp_complete_torus/original/original_z_mvp.ldl.zscore.torus.merged.csv"))
raw.ldl=raw.ldl[-which(raw.ldl$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]


df1=mash.ldl[,c(1:4)]
colnames(df1)=c("Outcome","log2OR","Lower","Upper")
df1$se=(df1$log2OR-df1$Lower)/1.96

df2=raw.ldl[,c(1:4)]

colnames(df2)=c("Outcome","log2OR","Lower","Upper")
df2$se=(df2$log2OR-df2$Lower)/1.96

# add a group column
df1$group <- "mash"
# create a second dataset, similar format to first

# different group
df2$group <- "raw"
# and we adjust the values a bit, so it will look different in the plot
df2[,c("log2OR","Lower","Upper")] 

# combine the two datasets                      
df = rbind(df1,df2)


z=(df1$log2OR-df2$log2OR)/sqrt(df1$se^2+df2$se^2)
p=2*pnorm(-1*abs(z))


dotCOLS = c("#a6d8f0","#f9b282")
barCOLS  <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
dotCOLS = c("blue2","red2")
barCOLS <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

df$p=c(p,p)
df$fp = format.pval(df$p,digits=1)

p = df %>% ggplot(aes(y=reorder(Outcome, desc(p)), x=log2OR, xmin=Lower, xmax=Upper, col=group, fill=group)) +
    geom_point(size=3, position=position_dodge(width = 0.5)) +
    geom_errorbarh(height=0, size=1, position=position_dodge(width = 0.5)) +
    geom_vline(xintercept=0, lty=2) +
    # geom_point(size=3, shape=21, colour="white", stroke = 0.5, position=position_dodge(width = 0.5)) +
    scale_fill_manual(values=barCOLS)+
    scale_color_manual(values=dotCOLS)+
    scale_y_discrete(name="Annotation") +
    scale_x_continuous(name="Log2 Odds ratio in ldl", limits = c(-5, 12), breaks=c(-5,0,5,10), labels=c(-5,0,5,"P-value"), position='top') +
    theme_minimal() +
    theme(legend.position=c(0.05,0.05), legend.justification=c(0,0), legend.title=element_blank()) +
    geom_text(aes(x=10, y=Outcome, label=fp), hjust=0.5, color='black')


p

```


###TG

```{r}
rm(list=ls())
library('ggplot2') 

mash.tg=read.csv(("~/Downloads/mvp_complete_torus//mash/tg.zscore.torus.merged.csv"))
mash.tg=mash.tg[-which(mash.tg$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]
###Orignal
raw.tg=read.csv(("~/Downloads/mvp_complete_torus/original/original_z_mvp.tg.zscore.torus.merged.csv"))
raw.tg=raw.tg[-which(raw.tg$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]


df1=mash.tg[,c(1:4)]
colnames(df1)=c("Outcome","log2OR","Lower","Upper")
df1$se=(df1$log2OR-df1$Lower)/1.96

df2=raw.tg[,c(1:4)]

colnames(df2)=c("Outcome","log2OR","Lower","Upper")
df2$se=(df2$log2OR-df2$Lower)/1.96

# add a group column
df1$group <- "mash"
# create a second dataset, similar format to first

# different group
df2$group <- "raw"
# and we adjust the values a bit, so it will look different in the plot
df2[,c("log2OR","Lower","Upper")] 

# combine the two datasets                      
df = rbind(df1,df2)


z=(df1$log2OR-df2$log2OR)/sqrt(df1$se^2+df2$se^2)
p=2*pnorm(-1*abs(z))




dotCOLS = c("#a6d8f0","#f9b282")
barCOLS = c("#008fd5","#de6b35")
dotCOLS = c("#a6d8f0","gray80")
barCOLS = c("#008fd5","gray80")

df$p=c(p,p)
df$fp = format.pval(df$p,digits=1)

p = df %>% ggplot(aes(y=reorder(Outcome, desc(p)), x=log2OR, xmin=Lower, xmax=Upper, col=group, fill=group)) +
    geom_point(size=3, position=position_dodge(width = 0.5)) +
    geom_errorbarh(height=0, size=1, position=position_dodge(width = 0.5)) +
    geom_vline(xintercept=0, lty=2) +
    # geom_point(size=3, shape=21, colour="white", stroke = 0.5, position=position_dodge(width = 0.5)) +
    scale_fill_manual(values=barCOLS)+
    scale_color_manual(values=dotCOLS)+
    scale_y_discrete(name="Annotation") +
    scale_x_continuous(name="Log2 Odds ratio in tg", limits = c(-5, 12), breaks=c(-5,0,5,10), labels=c(-5,0,5,"P-value"), position='top') +
    theme_minimal() +
    theme(legend.position=c(0.05,0.05), legend.justification=c(0,0), legend.title=element_blank()) +
    geom_text(aes(x=10, y=Outcome, label=fp), hjust=0.5, color='black')


p

```

###TC

```{r,echo=F}
rm(list=ls())
library('ggplot2') 

mash.tc=read.csv(("~/Downloads/mvp_complete_torus//mash/tc.zscore.torus.merged.csv"))
mash.tc=mash.tc[-which(mash.tc$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]
###Orignal
raw.tc=read.csv(("~/Downloads/mvp_complete_torus/original/original_z_mvp.tc.zscore.torus.merged.csv"))
raw.tc=raw.tc[-which(raw.tc$annotation=="GTEx_FE_META_TISSUE_GE_MaxCPP"),]


df1=mash.tc[,c(1:4)]
colnames(df1)=c("Outcome","log2OR","Lower","Upper")
df1$se=(df1$log2OR-df1$Lower)/1.96

df2=raw.tc[,c(1:4)]

colnames(df2)=c("Outcome","log2OR","Lower","Upper")
df2$se=(df2$log2OR-df2$Lower)/1.96

# add a group column
df1$group <- "mash"
# create a second dataset, similar format to first

# different group
df2$group <- "raw"
# and we adjust the values a bit, so it will look different in the plot
df2[,c("log2OR","Lower","Upper")] 

# combine the two datasets                      
df = rbind(df1,df2)


z=(df1$log2OR-df2$log2OR)/sqrt(df1$se^2+df2$se^2)
p=2*pnorm(-1*abs(z))


dotCOLS = c("#a6d8f0","#f9b282")
barCOLS = c("#008fd5","#de6b35")
dotCOLS = c("#a6d8f0","gray80")
barCOLS = c("#008fd5","gray80")

df$p=c(p,p)
df$fp = format.pval(df$p,digits=1)

p = df %>% ggplot(aes(y=reorder(Outcome, desc(p)), x=log2OR, xmin=Lower, xmax=Upper, col=group, fill=group)) +
    geom_point(size=3, position=position_dodge(width = 0.5)) +
    geom_errorbarh(height=0, size=1, position=position_dodge(width = 0.5)) +
    geom_vline(xintercept=0, lty=2) +
    # geom_point(size=3, shape=21, colour="white", stroke = 0.5, position=position_dodge(width = 0.5)) +
    scale_fill_manual(values=barCOLS)+
    scale_color_manual(values=dotCOLS)+
    scale_y_discrete(name="Annotation") +
    scale_x_continuous(name="Log2 Odds ratio in tc", limits = c(-5, 12), breaks=c(-5,0,5,10), labels=c(-5,0,5,"P-value"), position='top') +
    theme_minimal() +
    theme(legend.position=c(0.05,0.05), legend.justification=c(0,0), legend.title=element_blank()) +
    geom_text(aes(x=10, y=Outcome, label=fp), hjust=0.5, color='black')


p

```


