---
title: "mash_fit"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("mashr")
library("flashr")
```

# Fitting Mash Model
  
**Organize the data**

Here we provide with an overview of how the mash fit was conducted for a presorted set of summary statistics across the four lipid traits and provide the model fit object (`mfitMVP.rds`) so that users can recreate. 

**1) First, read in a matrix of summary statistics that have already been pruned to contain the same effect allele direction across phenotypes**

**2) Read in a pre-pruned matrix of maximum statistics across LD blocks**


```{r pressure, echo=FALSE,eval=FALSE}
zmash=readRDS("~/Dropbox/zmash.rds")
max_block=read.table("~/Dropbox/mash_materials/max_ld_block.txt")
source('~/Dropbox/mash_materials/flashscript.R')
```



**3) estimate VHat, the covariance of the errors, as the covariance matrix of the weakest effects**


```{r,eval=F}
# identify a random subset of 40000 tests from which to estimate weakest effects
random.subset = sample(1:nrow(znew),40000)
zmash=as.matrix(znew[,c("hdl","ldl","tg","tc")]);rownames(zmash)=ztab$snpid.x
data.temp = mash_set_data(zmash[random.subset,],alpha = 1)
Vhat = estimate_null_correlation_simple(data.temp)
#saveRDS(Vhat,".../data/MVPVhat.rds")
##Now add that to the mash object
data.random = mash_set_data(zmash[random.subset,],alpha = 1,V=Vhat)
```


**4) Estimate covariance matrices $U_{k}$**

```{r,eval=F}
zmax=apply(max_block[,c(4:7)],2,function(x){as.numeric(x)});rownames(zmax)=max_block$snpid.x
data.strong = mash_set_data(zmax,alpha = 1,V=Vhat)

U.pca = cov_pca(data.strong,3)

U.flash=cov_flash(data.strong, non_canonical = TRUE)
X.center = apply(data.strong$Bhat, 2, function(x) x - mean(x))
U.ed = cov_ed(data.strong, c(U.flash, U.pca, list("XX" = t(X.center) %*% X.center / nrow(X.center))))
U.c = cov_canonical(data.random)
```


## 5) Compute weights:

```{r,eval=F}
m = mash(data.random, Ulist = c(U.ed,U.c),outputlevel = 1)
saveRDS(m,"../data/mfitMVP.rds")
```



**6) Compute Posteriors**

```{r,eval=F}
for(i in 0:13){
  start=i*2e6+1
  stop=(i+1)*2e6
  print(c(start,stop))
  library("mashr")
  mash.data=mash_set_data(zmash[start:stop,],V = Vhat,alpha = 1)
  p=mash_compute_posterior_matrices(m$fitted_g, mash.data, algorithm.version = "Rcpp")
  saveRDS(p,file = paste0("../data//mash_mvp_batch",i,".rds"))
}
```

We have merged to create the shared files with posterior means (`combined_pm_MVP.rds`) and lfsr (`combined_lfsr_mvp.rds`).Because these are for the Exchangeable Z statistic model, the posterior means (E(b/shat|shat,Data)) are then rescaled by the standard error of beta hat to produce (E(b|shat,Data)): (`Merged_MVP_Full_se_raw`) to create `MVP_all_beta_posterior_beta` and `posterior_beta_se.txt` or the weighting algorithm in the PRS computations. 







