---
title: "prs_equivalency"
author: "Sarah Urbut"
date: "2021-09-11"
output: 
  workflowr::wflow_html:
    code_folding: hide
    toc: no
  html_notebook: default
  html_document:
    df_print: paged
    toc: no
editor_options:
  chunk_output_type: inline
---

## Introduction

We want to understand the equivalent quantile of PRS risk that equates with risk conveyed by traditional risk factors.

First, we organize the data:
```{r,eval=F}
rm(list=ls())
prs=read.table("~/Dropbox/cad_AnnoPred.profile",header=T)
cov=read.table("~/Dropbox/big_ukb_file.txt",header=T,sep="\t")
ukb_adj_cov=read.table("~/Dropbox/ukbb-lipids-meds.txt",sep = "\t",header = T)


i=as.character(intersect(intersect(prs$IID,cov$id),ukb_adj_cov$eid))

rownames(prs)=prs$IID
rownames(cov)=cov$id
rownames(ukb_adj_cov)=ukb_adj_cov$eid

df=na.omit(data.frame(prs[i,],"age"=cov[i,"age"],"sex"=cov[i,"Sex"],"cad"=cov[i,"Incd_Coronary_Artery_Disease_HARD"],cov[i,c("PC1","PC2","PC3","PC4","PC5")],"ldladj"=ukb_adj_cov[i,"ldladj"]))

df$SCORESUM=-1*df$SCORESUM

df$prs_quant=(df$SCORESUM-mean(df$SCORESUM))/sd(df$SCORESUM)
df$sex=as.factor(ifelse(df$sex=="Male",1,0))
df$ldl_fact=as.factor(ifelse(df$ldladj>190,1,0))
saveRDS(df,"~/Dropbox/df_ukb.rds")
```

```{r}
df=readRDS("~/Dropbox/df_ukb.rds")
head(df)
library(corrplot)
correlations <- cor(df[,c("ldladj","prs_quant","age","cad","PC1","PC2","PC3","PC4","PC5")])
corrplot(correlations, method="circle")
```

Note that SCORESUM is actually the opposite direction (increasing score is protective) so we take the opposite. Let's create a factor to call LDL greater or less than 190. Now, let's fit some simple logistic regression. Here's the Logistic with factor

```{r}
glm.fit=glm(df$cad~df$prs_quant+df$PC1+df$PC2+df$PC3+df$PC4+df$PC5+df$age+df$sex+df$ldl_fact,family=binomial)
exp(coef(glm.fit))

glm.ldl=glm(df$cad~df$ldl_fact,family=binomial)
exp(coef(glm.ldl))

glm.prs=glm(df$cad~df$prs_quant,family=binomial)
exp(coef(glm.prs))

s=sample(seq(1,length(df$prs_quant)),100000)
r=(rank(df$prs_quant)/length(df$prs_quant))

plot(df$prs_quant[s],exp(df$prs_quant[s]*coef(glm.prs)[[2]]),main="Conditional OR CAD for PRS_quantile",xlab='PRS_Quantile',ylab="Conditional OR")
abline(h=exp(coef(glm.ldl)[[2]]),col="red",pch=1)
abline(v=coef(glm.ldl)[[2]]/coef(glm.prs)[[2]],pch=2,col="green")
legend("topleft",c("OR_LDL>190","PRS_quantile"),col=c("red","green"),fill=c("red","green"))

q=coef(glm.ldl)[[2]]/coef(glm.prs)[[2]]

w=which(df$prs_quant>q)
t=which.min(df$prs_quant[w])
r[w[t]]
```

This means we'd need a PRS quantile of `r coef(glm.ldl)[[2]]/coef(glm.prs)[[2]]` which corresponds to the `r r[w[t]]` percentile!



## Predicting Risk

First consider discriminant analysis: how does the distirbution of CAD and LDL score discriminate between CAD and CAD free folks:
```{r}
# 
# par(mfrow=c(1,2))
# plot(density(r[df$cad==1]),col="red",main="PRS Quantile by CAD class",xlab="Quantile")
# lines(density(r[df$cad==0]),col="green")
# legend("bottomright",c("cad free","cad present"),col=c("green","red"),pch=1)

par(mfrow=c(1,3))
plot(density(df$prs_quant[df$cad==1]),col="red",main="PRS Z_Score Distribution by CAD class",xlab="Z_Score")
lines(density(df$prs_quant[df$cad==0]),col="green")
legend("bottomright",c("cad free","cad present"),col=c("green","red"),pch=1)

plot(density(df$ldladj[df$cad==0]),col="green",main="LDL_adj Distribution by CAD class",xlab="LDL-adjusted")
lines(density(df$ldladj[df$cad==1]),col="red")
legend("bottomright",c("cad free","cad present"),col=c("green","red"),pch=1)

zldl=df$ldladj-mean(df$ldladj)/sd(df$ldladj)
plot(density(zldl[df$cad==0]),col="green",main="Z_ldl by CAD",xlab="Z-Ldl")
lines(density(df$ldladj[df$cad==1]),col="red")
legend("bottomright",c("cad free","cad present"),col=c("green","red"),pch=1)
```
